The Efficient Handling of Guards in the Design of RPython's Tracing JIT

### 2012-10-16 talk/vmil2012/presentation/talk.tex
some tweaks to the text
add charts for guard failure rates combined for all benchmarks and for go

\usepackage[english]{babel}
\usepackage{listings}
\usepackage{ulem}
\usepackage{color}
\usepackage{alltt}

@@

\section{Introduction}

@@

\begin{frame}
  \frametitle{Guards}
  \begin{itemize}
      \item Points of control flow divergence are marked with guards
      \item Operations that check whether conditions are still true
      \item When a guard fails, execution of the trace stops and continues in the interpreter
      \pause
      \begin{block}{Guard Characteristics}
          \begin{itemize}
              \item lots of them, up to 20\% guards
              \item most never fail
              \item costly to implement
          \end{itemize}
      \end{block}
  \end{itemize}
\end{frame}

@@

\section{High-Level}

@@

\begin{frame}
  \frametitle{Interaction with Optimization}
  \begin{itemize}
      \item Some optimizations make it necessary to store extra information in symbolic frames
      \item examples:
          \begin{itemize}
              \item delayed heap stores (need to do stores before resuming interpreter)
              \item allocation removal (need to allocate objects before resuming)
          \end{itemize}
      \item can be compressed using similar techniques
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Emitting Guards}
  Guards are compiled as
  \begin{itemize}
    \item Quick Check if the condition holds
    \item and a mapping of machine locations to JIT-variables % indirection using the fail-boxes
  \end{itemize}
  In case of failure
  \begin{itemize}
    \item execution jumps to compensation code, decodes and stores mapping
    \item returns to interpreter that rebuilds state
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/loop.pdf}
  \label{fig:trampoline}
  \end{figure}
\end{frame}

\begin{frame}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/bridge_compiled.pdf}
  \label{fig:trampoline}
  \end{figure}
\end{frame}
\begin{frame}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/bridge_patched.pdf}
  \label{fig:trampoline}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Patching Guards for Bridges}
\end{frame}

\section{Evaluation}

>>>>

\usepackage[english]{babel}
\usepackage{listings}
\usepackage{ulem}
\usepackage{color}
\usepackage{alltt}
\usepackage{tikz}
\usepackage{pgfplots}

@@

%\section{Introduction}

@@

\begin{frame}
  \frametitle{Guards}
  \begin{itemize}
      \item Points of control flow divergence are marked with guards
      \item Operations that check whether conditions are still true
      \item When a guard fails, execution of the trace stops and continues in the interpreter
      \pause
      \item \emph{This talk:} describe technology and design decisions around guards
      \pause
      \begin{block}{Guard Characteristics}
          \begin{itemize}
              \item lots of them, up to 20\% guards
              \item most never fail
              \item need big information attached to them
          \end{itemize}
      \end{block}
  \end{itemize}
\end{frame}

@@

%\section{High-Level}

@@

\begin{frame}
  \frametitle{Interaction with Optimization}
  \begin{itemize}
      \item Some optimizations make it necessary to store extra information in symbolic frames
      \item examples:
          \begin{itemize}
              \item allocation removal (need to allocate objects before resuming)
              \item delayed heap stores (need to do stores before resuming interpreter)
          \end{itemize}
      \item can be compressed using similar techniques
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Emitting Guards}
  Guards are compiled as
  \begin{itemize}
    \item Quick Check if the condition holds
    \item and a mapping of machine locations to JIT-variables % indirection using the fail-boxes
  \end{itemize}
  In case of failure
  \begin{itemize}
    \item execution jumps to shared compensation code, decodes and stores mapping
    \item returns to interpreter that rebuilds state
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/loop.pdf}
  \end{figure}
\end{frame}

\begin{frame}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/bridge_compiled.pdf}
  \end{figure}
\end{frame}
\begin{frame}
  \begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/bridge_patched.pdf}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Patching Guards for Bridges}
\end{frame}
\begin{frame}[t,fragile]
\pgfplotsset{tick label style={font=\tiny\bfseries},
label style={font=\small},
legend style={font=\tiny}
}
    \frametitle{Guard Failure Rates}
    \begin{figure}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
            xlabel= Guards by failures,
            ylabel=Relative \# of failures,
            xtick=\empty,
            ytick=\empty,
            ]
                \input{figures/data}
            \end{axis}
        \end{tikzpicture}
    \end{figure}
\end{frame}
\begin{frame}[t,fragile]
\pgfplotsset{tick label style={font=\tiny\bfseries},
label style={font=\small},
legend style={font=\tiny}
}
    \frametitle{Guard Failure Rates / Go Benchmark}
    \begin{figure}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
            xlabel= Guards by failures,
            ylabel=Relative \# of failures,
            xtick=\empty,
            ytick=\empty,
            ]
                \input{figures/go_data}
            \end{axis}
        \end{tikzpicture}
    \end{figure}
\end{frame}

%\section{Evaluation}
